FlowFPX: Nimble Tools for Debugging Floating-Point Exceptions

### 2023-08-04 paper.tex
Cleanup section 4

\begin{description}
\item[\texttt{active::Bool}] Inject only if this is true.
\item[\texttt{n\_inject::Int64}] Inject at most this many \NaN{}s.
\item[\texttt{odds::Int64}] Inject if \texttt{rand(odds) == 1}.
\item[\texttt{functions::Array}] Inject if we are within the dynamic extent of any of these functions.
\item[\texttt{libraries::Array}] Inject if the top-most stack frame\footnote{After stripping off the stack frames that come from \FT{} itself.} comes from any of these libraries.
\item[\texttt{replay::String}] Name of a file to read an injection recording from.
\item[\texttt{record::String}] Name of a file to write injection events to for later replay.
\end{description}

@@

\subsubsection{Graph diffing}

CSTG can diff two graphs to show how the flows changed.
For example, we took the \FT{} logs from running ShallowWaters\cite{klowerNumberFormatsError2020} \footnote{\Cref{s:sw}}, and split the logs at the 10\% mark.
We used CSTG to diff the flow graphs between the first 10\% of the run of the program and the remaining 90\%.
The diff can be seen in \cref{fig:cstg_diff_demo}.

@@

\FlowFPX{} is effective for debugging all sorts of \NaN{}-related issues.
We have used it to reveal silent \NaN{}s, uncover the root cause of \NaN{}s,
and fuzz for weak points in several case studies.

\subsection{Surprise \NaN{}s: ShallowWaters}
\label{s:sw}

Most physics simulations of this sort include a parameter called the Courant-Friedrichs-Lewy (``CFL'') parameter, which roughly describes the size of the time step to take in running the simulation.
Normally this value sits between 0 and 1---smaller values mean the simulation runs slower but with higher accuracy, whereas higher values mean the simulation takes bigger time steps and consequently runs faster, but results might be inaccurate.
The primary reason for problems arising comes from information not having enough time to propagate through the simulation before the next time step is taken.

@@

For our first case study, we looked at the \texttt{ShallowWaters} library~\cite{klowerNumberFormatsError2020,klowerPositsAlternativeFloats2019,klowerLowprecisionClimateComputing2021}, which runs a simulates the flow of water over a seabed.
This is not a general shallow waters simulation library---it is designed primarily to showcase numerical methods involving 16-bit floats.
That said, it works for our purposes as a proving ground for FloatTracker.
We dialed the CFL parameter up to high values (as high as 10) and we started seeing \NaN{}s in the result.
Note that the program did not crash: instead, it began producing bad output.

@@

Running FloatTracker can produce a lot of logs.
To get a quick overview of the most common paths to problems, we can use CSTG to get an overview of what stack frames appear the most often.
% FIXME: cite CSTG

@@

\subsection{\NaN{} fuzzing to find a kill: OrdinaryDiffEq}

@@

\subsection{A Bug-Hunting Safari}

% Notes on RxInfer

We searched GitHub for open issues in Julia libraries involving difficult-to-find \NaN{}s.
In our first batch of issues, we found the \code{RxInfer.jl} library with an issue revolving around \NaN{} detection.\footnote{\url{https://github.com/biaslab/RxInfer.jl/issues/116}}
We offered our tool for the package authors to use, and within a day the authors were able to track down the location of a \NaN{} and fix the issue---all with minimal guidance.

@@

>>>>

\begin{itemize}
\item \textbf{\texttt{active::Bool} Inject only if this is true.
\item \texttt{n\_inject::Int64} Inject at most this many \NaN{}s.
\item \texttt{odds::Int64} Inject if \texttt{rand(odds) == 1}.
\item \texttt{functions::Array} Inject if we are within the dynamic extent of any of these functions.
\item \texttt{libraries::Array} Inject if the top-most stack frame\footnote{After stripping off the stack frames that come from \FT{} itself.} comes from any of these libraries.
\item \texttt{replay::String} Name of a file to read an injection recording from.
\item \texttt{record::String} Name of a file to write injection events to for later replay.
\end{itemize}

@@

@@

We have used \FlowFPX{} to debug and harden various programs against exceptional values, which we now present in the following case studies.

\begin{itemize}
\item \Cref{s:sw} uses the \FT{} and CSTG to uncover the source of \NaN{} and \Inf{} in a simulation.
\item \Cref{s:ode} uses the fuzzing abilities of \FT{} to harden an existing library against unexpected exceptional values.
\item \Cref{s:finch} is another case where fuzzing uncovered a hard-to-find bug.
\item \Cref{s:ocean} showcases using \GPUFPX{} in a simulation.
\end{itemize}

Besides case studies that we ourselves did, \cref{s:safari} is an ``in the wild'' usage of \FT{} that helped researchers uncover the source of a \NaN{} in their own code.

\subsection{ShallowWaters}
\label{s:sw}

For our first case study, we looked at the \texttt{ShallowWaters} library~\cite{klowerNumberFormatsError2020,klowerPositsAlternativeFloats2019,klowerLowprecisionClimateComputing2021}, which runs a simulates the flow of water over a seabed.
This is not a general shallow waters simulation library---it is designed primarily to showcase numerical methods involving 16-bit floats.
That said, it works for our purposes as a proving ground for FloatTracker.

Many physics simulations of this sort include a parameter called the Courant-Friedrichs-Lewy (``CFL'') parameter, which roughly describes the size of the time step to take in running the simulation.
Normally this value sits between 0 and 1---smaller values mean the simulation runs slower but with higher accuracy, whereas higher values mean the simulation takes bigger time steps and consequently runs faster, but results might be inaccurate.
The primary reason for problems arising comes from information not having enough time to propagate through the simulation before the next time step is taken.

@@

We dialed the CFL parameter up to extraordinarily high values and we started seeing \NaN{}s in the result.
\code{ShallowWaters} detected the instability, but not before producing some bad output.

@@

Running FloatTracker can produce a lot of logs.
To get a quick overview of the most common paths to problems, we can use CSTG to get an overview of what stack frames appear the most often.

@@

Another one of CSTGs abilities is \emph{graph diffing}.
We took the \FT{} logs from running ShallowWaters and split the logs at the 10\% mark.
We used CSTG to diff the flow graphs between the first 10\% of the run of the program and the remaining 90\%.
The diff can be seen in \cref{fig:cstg_diff_demo}.

\subsection{\NaN{} fuzzing to find a kill: OrdinaryDiffEq}
\label{s:ode}

@@

@@

\subsection{A Bug-Hunting Safari}
\ref{s:safari}

\FT{} has proven to be useful for more than its creators---we discovered an open issue in the \code{RxInfer.jl} library with an issue revolving around \NaN{} detection.\footnote{\url{https://github.com/biaslab/RxInfer.jl/issues/116}}
Even though we were unable to provide hands-on guidance as the code in question was proprietary, the authors of \code{RxInfer.jl} were able use \FT{} themselves, and within a day they tracked down the location of a \NaN{} gen.
The \code{RxInfer.jl} authors provided us with valuable feedback on \FT{}'s API.
